IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-08 paper/intro.tex

\section{Introduction}
\label{sec:intro}
\Red{DM had some very good comments on how to make this sound less
  like a pitch for LIO. Outline for something more compelling (not
  necessarily in this order):
\begin{itemize}
  \item Motivate IFC
  \item Coarse grained IFC is easy
  \item Coarse grained IFC is sufficient (vs. fine grained); do by citation
  \item We have the right API and it's language-agnostic; proof is also language independent
  \item You can ``drop it in''
\end{itemize}
}

@@

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve the language JavaScript, and retrofit dynamic information
flow control on top of it.

Systems that track information flow at a fine-grained level, e.g. per
value in the system, offer great flexibility.  However, they
require deep integration with the programming language and modify
the semantics in non-trivial ways.  This makes fine-grained systems
less suitable in contexts where preserving an existing language
is important.
%
Moreover, the task of adding fine-grained IFC even to a simple language is
non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
such as JavaScript, this challenge becomes almost insurmountable.

@@

A further benefit of the coarse-grained approach
is that a single IFC design and semantics can be generalized
to a large class of programming languages.
Leveraging this insight, we define a form of coarse-grained
dynamic IFC and show how it can be combined with any programming
language for which external effects can be controlled.
At the core of our IFC system is a sandboxing construct that isolates
potentially sensitive computations, allowing them to
communicate only through clearly specified interfaces.
Using an approach by Matthews-Findler~\cite{Matthews:2007:OSM:1190216.1190220}
for combining operational semantics,
we formally state and prove non-interference guarantees that are
independent of the choice of specific target language.

One challenge in the design of our system is that the semantics for the
retrofitted IFC language may not accurately correspond to the optimized
implementation details of a particular target language.
We discuss and partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then non-interference in the abstract specification
carries over to the concrete implementation.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  information flow control language free of non-IFC constructs.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Section~\ref{sec:retrofit})
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC language that preserves non-interference. (Section~\ref{sec:formal}
  and~\ref{sec:concrete})
  \item We have implemented an IFC system based on these semantics
  for Node.js, and we connect our formalism to another implementation
  based on this work
  for client-side JavaScript in Gecko~\cite{swapi}.
  Furthermore, we outline an implementation for the C programming
  language.
  (Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}

>>>>

\section{Introduction}
\label{sec:intro}
%   \Red{DM had some very good comments on how to make this sound less
%     like a pitch for LIO. Outline for something more compelling (not
%     necessarily in this order):
%   \begin{itemize}
%     \item Motivate IFC
%     \item Coarse grained IFC is easy
%     \item Coarse grained IFC is sufficient (vs. fine grained); do by citation
%     \item We have the right API and it's language-agnostic; proof is also language independent
%     \item You can ``drop it in''
%   \end{itemize}
%   }

@@

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve JavaScript's semantics while retrofitting it with dynamic information
flow control.

One approach to information flow control is fine-grained tracking,
which offers great flexibility.
However, these systems
require deep integration with the programming language and modify
the semantics in non-trivial ways.  This makes fine-grained systems
less suitable in contexts where preserving an existing language
is important.
%
Moreover, the task of adding fine-grained IFC even to a simple language is
non-trivial~\cite{hritcu2013testing}; when considering a real-world language,
such as JavaScript, this challenge becomes almost insurmountable.

@@

In this paper, we take coarse-grained labeling~\cite{Zeldovich:2006,
lio} to its logical extreme: an information flow control system should
be thought of as multiple instances of completely isolated language
runtimes or \emph{tasks}, with information flow control applied to
inter-task communication.  We describe a formal system in which an IFC
system can be designed once and then applied to any programming language
which has control over external effects (e.g., JavaScript or C with
access to hardware privilege separation).  We formalize this system
using an approach by
Matthews-Findler~\cite{Matthews:2007:OSM:1190216.1190220} for combining
operational semantics and prove non-interference guarantees
that are independent of the choice of specific target language.

What distinguishes this setting from previous coarse-grained IFC
systems?
%
First, even though the underlying semantic model involves communicating
tasks, these tasks can be coordinated together in ways that simulate
traditional features in monolithic languages.  In fact, simulating
features in this way is a very useful \emph{design tool} for discovering
what variants of the features are permissible and which are not.
%
Second, although completely separate tasks are semantically easy to
reason about, real-world implementations often blur the lines between
tasks in the name of efficiency.
Characterizing what optimizations are permissible is subtle, since
removing transitions from the operational semantics of a language can
break non-interference.  We partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then non-interference in the abstract specification
carries over to the concrete implementation.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  dynamic information flow control language free of non-IFC constructs.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Section~\ref{sec:retrofit})
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC language that preserves non-interference. (Section~\ref{sec:formal}
  and~\ref{sec:concrete})
  \item We have implemented an IFC system based on these semantics
  for Node.js, and we connect our formalism to another implementation
  based on this work
  for client-side JavaScript in Gecko~\cite{swapi}.
  Furthermore, we outline an implementation for the C programming
  language.
  (Section~\ref{sec:real})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, labeled references, or a
  notion of clearance. (Section~\ref{sec:extensions})
\end{itemize}

### 2014-07-08 paper/retrofit.tex
wip, pulling

Information flow control systems operate by associating data with \emph{labels},
and specifying whether or not data tagged with one label |l1| can flow
to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), and the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{non-interference}.

Supporting fine-grained labeling of data in a programming language often
requires extremely intrusive modifications to its semantics and
implementation.  In this paper, we advocate applying coarse-grained
labeling~\cite{Zeldovich:2006, lio} at its most extreme: our starting point
is multiple instances of completely isolated language runtimes or \emph{tasks}, with
information flow control applied between their communication.  For
languages which are implemented with isolation, such as JavaScript or C
with access to hardware privilege separation, systems along these lines
are far more easily implementable and require fewer semantic changes.

What distinguishes this setting from previous coarse-grained IFC
systems?
%
First, even though the underlying semantic model involves communicating
tasks, these tasks can be coordinated together in ways that simulate
traditional features in monolithic languages.  In fact, simulating
features in this way is a very useful \emph{design tool} for discovering
what variants of the features are permissible and which are not.
%
Second, although completely separate tasks are semantically easy to
reason about, real-world implementations often blur the lines between
tasks in the name of efficiency (Section~\ref{sec:concrete}).
Characterizing what optimizations are permissible is subtle, since
removing transitions from the operational semantics of a language can
break non-interference.

To give a simple example, here is a low security task sending a high security
task a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) l (IT (ref true)))|
\end{align*}
For now, ignore the tags |TI| and |IT|: roughly, this code creates a new
|sandbox|ed task with identifier $i$ which waits (\textbf{blockingRecv}) for a
message, and then \textbf{send}s the task a mutable reference |(ref
true)| which is labeled $l$.  If this operation actually shared the mutable cell, this cell
could be used to violate information flow control, by leaking
information from the high security task to the low security task.

This example demonstrates both of our design principles.  When working
with isolated tasks, each of which have their own heaps, the isolation
is obviously true: the sandboxed task receives a dangling pointer.  From
there, we can further ask: given the obviously safe formulation with
separate heaps, can this fact be useful for showing the safety of a
semantics which only has a single heap (we answer this question
positively in Section~\ref{sec:single-heap})?  Finally, this na\"ive
semantics violates expectations about how inter-task communication works
on a shared heap, and strongly suggests one should restrict the types of
data which may be passed between tasks (for example, in JavaScript, one
might only allow JSON objects to be passed between tasks, rather than
general object structures).

Before moving on to the formal treatment of our system, we'd like to give
two more examples of language features which have an interesting interpretation
when viewed in the setting of multiple tasks.

\paragraph{Stack frames versus scheduling}  One feature which is
supported by some information flow control systems is the ability to
temporarily raise the current label in order to do computation on data
(as done by LIO's |toLabeled| and Breeze's
brackets~\cite{Hritcu:2013:YIB:2497621.2498098,
stefan:2012:arxiv-flexible}).  After the computation, the label is lowered back to the original label,
enforcing a stack discipline.  This is the trickiest part of the feature: care must be taken to ensure that information doesn't leak when
the label is lowered. In fact, systems with this feature generally only
achieve termination-insensitive non-interference: if the high secrecy
section doesn't terminate, this is observable by the non-execution of
the defunct public continuation.

@@

When stack frames are treated as isolated tasks, the correct semantics
are achieved automatically.
%
Exceptions simply stop when the hit the top-level of a task:
the only way for the frame to communicate information is by
initiating inter-task message passing.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.
\Red{TODO: note that if the target language allows inspecting if
  result is exception we can implement Breeze's NAV (vs. just
propagating)}

@@

As mentioned previously, most modern, dynamic information flow control
languages encode policy by associating a label with data.  Our
embedding is agnostic to the choice of labeling scheme; we only require
the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056} with
the partial order $\sqsubseteq$, join |lub|, and meet |glb|.  In this
paper, we simply represent labels with the metavariable $l$, but do not
discuss them in more detail.

In our coarse-grained floating-label approach, labels are associated with tasks.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.
%
Hence before performing a read or a write, the IFC monitor inspects the
current label to decide whether the operation is permitted.
%
A task can only write to entities that are at least as sensitive.
%
Similarly, it can only read from entities that are less sensitive.
%
However, as in other floating-label systems, this current label can be raised
to allow the task to read from more sensitive entities at the cost of giving up
the ability to write to certain entities.

@@

These asynchronous communication primitives are worth further
discussion.  When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Messages can only be received by a task if its current label is
at least as high as the label of the message.
Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the message and the senders task identifier
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|, and otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more sensitive
than their current label would allow.

>>>>

Before moving on to the formal treatment of our system, we'd like to give
a brief primer of information flow control and describe some example programs
in our system, emphasizing the parallel between their implementation
in a multi-task setting, and the traditional, ``monolithic'' programming
language feature they simulate.

Information flow control systems operate by associating data with \emph{labels},
and specifying whether or not data tagged with one label |l1| can flow
to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), and the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{non-interference}.
%
In our coarse-grained floating-label approach, labels are associated with tasks.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.

For our first example, here is a program which spawns a new isolated task,
and then sends it a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) l (IT (ref true)))|
\end{align*}
For now, ignore the tags |TI| and |IT|: roughly, this code creates a new
|sandbox|ed task with identifier $i$ which waits (\textbf{blockingRecv}, binding $x$ with the received message) for a
message, and then \textbf{send}s the task a mutable reference |(ref
true)| which it labels $l$.  If this operation actually shared the mutable cell between the two tasks, it
could be used to violate information flow control if the tasks had differing labels.

While simple, this example demonstrates some of our general design principles.  When these
constructs are treated as isolated tasks, each of which have their own heaps, it
is obviously the case that there is no sharing: in fact, the sandboxed task receives a dangling pointer.  From
there, we can further ask: given the obviously safe formulation with
separate heaps, can this fact be useful for showing the safety of a
semantics which only has a single heap? (We answer this question
positively in Section~\ref{sec:single-heap}.)  Finally, this na\"ive
semantics violates expectations about how inter-task communication works
on a shared heap and strongly suggests one should restrict the types of
data which may be passed between tasks (for example, in JavaScript, one
might only allow JSON objects to be passed between tasks, rather than
general object structures).

Here are two more examples of language features which have an
interesting interpretation when viewed in the setting of multiple tasks.

\paragraph{Stack frames versus scheduling}  One feature which is
supported by some information flow control systems is the ability to
temporarily raise the current label in order to do computation on data
(as done by LIO's |toLabeled| and Breeze's
brackets~\cite{Hritcu:2013:YIB:2497621.2498098,
stefan:2012:arxiv-flexible}).  After the computation, the label is lowered back to the original label,
enforcing a stack discipline.  When this occurs, care must be taken to ensure that information doesn't leak when
the label is lowered (e.g., the result must not be readable). In fact, systems with this feature generally only
achieve termination-insensitive non-interference: if the high secrecy
section doesn't terminate, this is observable by the non-execution of
the defunct public continuation.

@@

When stack frames are treated as isolated tasks, the correct semantics
are achieved automatically.
%
Exceptions simply stop when the hit the top-level of a task:
the only way for the frame to communicate information is by
initiating inter-task message passing.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.

@@

As mentioned previously, most modern, dynamic information flow control
languages encode policy by associating a label with data.  Our
embedding is agnostic to the choice of labeling scheme; we only require
the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056} with
the partial order $\sqsubseteq$, join |lub|, and meet |glb|.  In this
paper, we simply represent labels with the metavariable $l$, but do not
discuss them in more detail.
%
To enforce labels, the IFC monitor inspects the
current label  before performing a read or a write to decide whether the operation is permitted.
%
A task can only write to entities that are at least as sensitive.
%
Similarly, it can only read from entities that are less sensitive.
%
However, as in other floating-label systems, this current label can be raised
to allow the task to read from more sensitive entities at the cost of giving up
the ability to write to certain entities.

@@

These asynchronous communication primitives are worth further
discussion.  When a task is sending a message using |send|, it also labels that
message with a label |il'| (which must be above the tasks current label |il|).
Messages can only be received by a task if its current label is
at least as high as the label of the message.
Specifically, receiving a message using
\[ |recvf ix1 ix2 ie1 ie2| \]
binds the message and the sender's task identifier
to local variables |ix1| and |ix2|, respectively, and then executes |ie1|.
Otherwise the message gets removed from the queue by rule \textsc{I-noRecv},
and that tasks continues its execution with |ie2|.
We denote the filtering of the message queue by |filter il Q|,
which is defined as follows.
If |Q| is the empty list |nil|, the
function is simply the identity function, i.e.,
|filter il nil = nil|, and otherwise:
\[
|filter il ((il', id, ie) , Q )| = \left\{
\begin{array}{l l}
|(il', id, ie) , filter il Q| & \quad \text{if |il' canFlowTo il|}\\
|filter il Q| & \quad \text{otherwise}
\end{array} \right.
\]
This ensures that tasks cannot receive messages that are more sensitive
than their current label would allow.
